---
title: "Palmer Penguins Analysis with R"
author: "Laith Saeed"
date: "2026-09-28"
---

## Exploring the Palmer Penguins Dataset

This post explores differences in body mass and flipper length among penguin species using R.

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER.



```{r}
#| message: false
#| warning: false

library(palmerpenguins)
library(ggplot2)
library(dplyr)
```


### EDA of Dataset and Aggregating
```{r}
penguins |> 
    head()
```



```{r}
species_summary <- penguins |>
   dplyr :: group_by(species) |>
   dplyr :: summarise(
        avg_body_mass = mean(body_mass_g, na.rm = TRUE),
        avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
    )

species_summary
```

### Body Mass by Species



```{r}
#| label: fig-r-body-mass
#| fig-cap: "Average body mass of Palmer penguins by species."

ggplot(species_summary, aes(x = species, y = avg_body_mass)) +
    geom_col() +
    labs(
        x = "Species",
        y = "Average Body Mass (g)",
        title = "Average Body Mass by Species"
    )
```
Genboo is heaviest on average then tie between Chinstrap and Adelie.


### Flipper Length and Body Mass


```{r}
#| label: fig-r-flipper-mass
#| fig-cap: "Relationship between flipper length and body mass in Palmer penguins."

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
    geom_point() +
    labs(
        x = "Flipper Length (mm)",
        y = "Body Mass (g)",
        title = "Flipper Length vs Body Mass"
    )
```

The scatterplot shows a positive association between body mass and flipper length. Penguins with longer flippers generally have greater body mass.


