---
title: "Palmer Penguins with R and Python"
author: "Laith Saeed"
date: "2026-09-28"
---

## Combining R and Python

This post demonstrates interoperability between R and Python using `reticulate`. R is used to create a summary of the Palmer Penguins dataset, and the resulting object is passed to Python for further analysis.

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER.



```{r}
#| message: false
#| warning: false

library(palmerpenguins)
library(dplyr)
library(reticulate)
```



```{r}
species_summary <- penguins |>
    group_by(species) |>
    summarise(
        avg_body_mass = mean(body_mass_g, na.rm = TRUE),
        avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
    )

species_summary
```

## Passing the R Summary to Python

The `species_summary` object was created in R. The following Python code accesses this R object using `reticulate` and uses it for further analysis.


```{python}
import matplotlib.pyplot as plt

summary_python = r.species_summary

summary_python
```

```{python}
#| label: fig-reticulate-body-mass
#| fig-cap: "Average body mass plotted in Python using summary data created in R."

plt.bar(
    summary_python["species"],
    summary_python["avg_body_mass"]
)

plt.xlabel("Species")
plt.ylabel("Average Body Mass (g)")
plt.title("R Summary Visualized with Python")
plt.show()
```

```{r}
#| label: fig-r-flipper-mass
#| fig-cap: "Relationship between flipper length and body mass in Palmer penguins."

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
    geom_point() +
    labs(
        x = "Flipper Length (mm)",
        y = "Body Mass (g)",
        title = "Flipper Length vs Body Mass"
    )

    The scatterplot shows a positive association between body mass and flipper length. Penguins with longer flippers generally have greater body mass.